---
title: "Bayesian Test for Homoscedasticity - Results Analysis"
output:
  html_document: default
  pdf_document: default
date: "2025-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r - packages}
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(BAS)
library(MASS)
library(BayesFactor)
library(stringr)
library(rstanarm)
library(quantreg)
```

### Introduction Graph
```{r - Introduction Graph}
set.seed(123)
n <- 300 
x <- seq(0, 1, length.out = n)

df <- rbind(
  data.frame(x = x, residual = res_homo, type = "Homoscedastic"),
  data.frame(x = x, residual = res_hetero, type = "Heteroscedastic")
)

# Set facet order
df$type <- factor(
  df$type,
  levels = c("Homoscedastic", "Heteroscedastic")
)

p_resid <- ggplot(df, aes(x = x, y = residual)) +
  geom_point(
    shape = 16,
    size  = 1.6,
    alpha = 0.55,
    colour = "black"
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  facet_wrap(~ type, nrow = 1) +
  labs(
    x = "Predictor",
    y = "Residuals"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8, colour = "black"),
    axis.ticks = element_line(linewidth = 0.8, colour = "black"),
    axis.ticks.length = unit(3.5, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 13),
    panel.spacing.x = unit(10, "pt"),
    panel.grid = element_blank()
  )

print(p_resid)
```

#   Bayesian Test for Homoscedasticity - Main Function 
```{r - Homoscedasticity Function}
bayesian_homoscedasticity_test <- function(
    formula,
    data,
    account_uncertainty = FALSE,
    S = 500,
    priors = c(H0 = 1/3, H1 = 1/3, H2 = 1/3),
    prior_reg = rstanarm::normal(0, 10),
    evidence = c("JZS", "BIC"),
    rscale = "medium"
) {

  evidence <- match.arg(evidence)

  # ---- Prior checks ----
  if (length(priors) != 3)
    stop("priors must be a named vector: c(H0=..., H1=..., H2=...)")

  if (!all(c("H0", "H1", "H2") %in% names(priors)))
    stop("priors must be named c(H0=..., H1=..., H2=...)")

  priors <- priors[c("H0", "H1", "H2")]
  priors <- priors / sum(priors)

  # ---- Stage 1: Mean model ----
  fit <- rstanarm::stan_glm(
    formula,
    data = data,
    prior = prior_reg,
    chains = 4,
    iter = 4000,
    refresh = 0,
    adapt_delta = 0.95
  )

  mm <- model.matrix(formula, data = data)
  y  <- model.response(model.frame(formula, data = data))

  if (ncol(mm) < 2)
    stop("The model must include at least one predictor.")

  x <- mm[, 2]
  x_center <- scale(x, center = TRUE, scale = FALSE)

  # ---- Function for posterior probabilities ----
  compute_probs <- function(y2) {

    df2 <- data.frame(
      y2 = y2,
      x = x_center,
      x2 = x_center^2
    )

    if (evidence == "BIC") {

      bic0 <- BIC(lm(y2 ~ 1,  data = df2))
      bic1 <- BIC(lm(y2 ~ x,  data = df2))
      bic2 <- BIC(lm(y2 ~ x2, data = df2))

      log_m <- c(
        H0 = -0.5 * bic0,
        H1 = -0.5 * bic1,
        H2 = -0.5 * bic2
      )

    } else {

      bf1 <- BayesFactor::lmBF(
        y2 ~ x,
        data = df2,
        rscaleCont = rscale
      )

      BF10 <- as.numeric(
        BayesFactor::extractBF(bf1, logbf = FALSE)[1, "bf"]
      )

      bf2 <- BayesFactor::lmBF(
        y2 ~ x2,
        data = df2,
        rscaleCont = rscale
      )

      BF20 <- as.numeric(
        BayesFactor::extractBF(bf2, logbf = FALSE)[1, "bf"]
      )

      if (!is.finite(BF10) || !is.finite(BF20) ||
          BF10 <= 0 || BF20 <= 0) {
        return(c(H0 = NA_real_, H1 = NA_real_, H2 = NA_real_))
      }

      log_m <- c(
        H0 = 0,
        H1 = log(BF10),
        H2 = log(BF20)
      )
    }

    log_prior <- log(priors[c("H0", "H1", "H2")])
    log_num <- log_prior + log_m

    w <- exp(log_num - max(log_num))
    post <- w / sum(w)

    post
  }

  # ---- No uncertainty version ----
  if (!account_uncertainty) {

    yhat <- as.vector(
      rstanarm::posterior_linpred(
        fit,
        transform = TRUE,
        draws = 1000
      ) %>%
        colMeans()
    )

    res <- y - yhat
    y2 <- res^2

    post <- compute_probs(y2)

    draws_out <- NULL
    n_skipped <- NULL

  } else {

    # ---- Uncertainty propagation ----
    draws <- as.matrix(fit)
    beta_draws <- draws[, colnames(mm), drop = FALSE]
    n_post <- nrow(beta_draws)

    pick <- sample.int(n_post, S, replace = TRUE)

    post_mat <- matrix(NA_real_, nrow = S, ncol = 3)
    colnames(post_mat) <- c("H0", "H1", "H2")

    for (s in seq_len(S)) {

      beta_s <- beta_draws[pick[s], ]

      yhat_s <- as.vector(mm %*% beta_s)

      res_s <- y - yhat_s

      y2_s <- res_s^2

      post_mat[s, ] <- compute_probs(y2_s)
    }

    post <- colMeans(post_mat, na.rm = TRUE)

    draws_out <- as.data.frame(post_mat)

    n_skipped <- sum(!is.finite(post_mat[, "H0"]))
  }

  # ---- Final output ----
  postH0 <- unname(post["H0"])
  postH1 <- unname(post["H1"])
  postH2 <- unname(post["H2"])

  best_n <- names(which.max(c(
    H0 = postH0,
    H1 = postH1,
    H2 = postH2
  )))

  interpretation <- c(
    H0 = "Evidence supports homoscedasticity.",
    H1 = "Evidence supports linear heteroscedasticity.",
    H2 = "Evidence supports nonlinear heteroscedasticity."
  )[best_n]

  out <- list(
    posterior_prob_H0 = postH0,
    posterior_prob_H1 = postH1,
    posterior_prob_H2 = postH2,
    account_uncertainty = account_uncertainty,
    S = if (account_uncertainty) S else NULL,
    priors = priors,
    evidence = evidence,
    rscale = if (evidence == "JZS") rscale else NULL,
    draws = draws_out,
    n_skipped = n_skipped,
    interpretation = interpretation
  )

  class(out) <- "results_simple"

  out
}


print.results_simple <- function(x, ...) {

  cat("Bayesian Test for Homoscedasticity\n")

  cat(sprintf(
    "Posterior prob (H0 - homoscedasticity)              = %.4f\n",
    x$posterior_prob_H0
  ))

  cat(sprintf(
    "Posterior prob (H1 - linear heteroscedasticity)     = %.4f\n",
    x$posterior_prob_H1
  ))

  cat(sprintf(
    "Posterior prob (H2 - nonlinear heteroscedasticity)  = %.4f\n",
    x$posterior_prob_H2
  ))


  cat("\n")
  cat(x$interpretation, "\n")
}
```

# Results Section Analyses

## 3.1 Simulation Study: Baseline Performance 
### 3.1.1 Homoscedastic Data
```{r - Homoscedastic Data}
# We generate a dataset following a homoscedastic structure
set.seed(123)
n <- 100 
x_homosc <- rnorm(n)
y_homosc <- 1 + 2 * x_homosc + rnorm(n, sd = 1)
df_homosc <- data.frame(x=x_homosc,y=y_homosc)

# We fit a linear model
lm <- lm(y_homosc ~ x_homosc)
pred_y <- predict(lm)
residuals <- resid(lm)

# Plot 1: Variance Structure
plot1 <- ggplot(data.frame(x_homosc, y_homosc), aes(x = x_homosc, y = y_homosc)) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    linewidth = 0.9,
    colour = "grey30"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.5,
    colour = "black"
  ) +
  labs(
    x = "Predictor",
    y = "Response"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )

print(plot1)

# Plot 2: Residuals vs Fitted values
plot2 <- ggplot(
  data.frame(Fitted = pred_y, Residuals = residuals),
  aes(x = Fitted, y = Residuals)
) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.55,
    colour = "black"
  ) +
  labs(
    x = "Fitted values",
    y = "Residuals"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )

print(plot2)

# TEST
bayesian_homoscedasticity_test(formula = y_homosc ~ x_homosc, data = df_homosc)
```
### 3.1.2 Linear Heteroscedastic Data
```{r - Linear Heteroscedasticity}
# We generate a dataset following a linear heteroscedastic structure
set.seed(123)
n <- 100                        
x_het_lin <- runif(n, 0, 10)    
sigma <- 1 + 0.5 * x_het_lin         
y_het_lin <- 1 + 2 * x_het_lin + rnorm(n, sd = sigma)
df_het_lin <- data.frame(x_het_lin = x_het_lin, y_het_lin = y_het_lin)

# We fit a linear model
lm_het_lin         <- lm(y_het_lin ~ x_het_lin, data = df_het_lin)
pred_y_het_lin     <- predict(lm_het_lin)
residuals_het_lin  <- resid(lm_het_lin)

# PLOT 1: Variance Structure
plot1_het_lin <- ggplot(
  df_het_lin,
  aes(x = x_het_lin, y = y_het_lin)
) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    linewidth = 0.9,
    colour = "grey30"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.5,
    colour = "black"
  ) +
  labs(
    x = "Predictor",
    y = "Response"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )

print(plot1_het_lin)

# PLOT 2: Residuals vs Fitted Values
plot2_het_lin <- ggplot(
  data.frame(
    Fitted    = pred_y_het_lin,
    Residuals = residuals_het_lin
  ),
  aes(x = Fitted, y = Residuals)
) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.55,
    colour = "black"
  ) +
  labs(
    x = "Fitted values",
    y = "Residuals"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )
print(plot2_het_lin)

#TEST
bayesian_homoscedasticity_test(formula = y_het_lin ~ x_het_lin, data = df_het_lin)
```

### 3.1.3 Nonlinear Heteroscedastic Data
```{r - Nonlinear Heteroscedasticity}
set.seed(13)
n_het_nl       <- 100                             
x_het_nl       <- runif(n_het_nl, 0, 10)         
sigma_het_nl   <- 1 + 0.2 * (x_het_nl - mean(x_het_nl))^2           
y_het_nl       <- 1 + 2 * x_het_nl + 
                  rnorm(n_het_nl, sd = sigma_het_nl)

df_het_nl <- data.frame(x_het_nl = x_het_nl, y_het_nl = y_het_nl)

# Fit a nonlinear model
lm_het_nl         <- lm(y_het_nl ~ x_het_nl, data = df_het_nl)
pred_y_het_nl     <- predict(lm_het_nl)
residuals_het_nl  <- resid(lm_het_nl)

# Checking the variance structure
# Plot 1: Variance Structure
plot1_het_nl <- ggplot(
  df_het_nl,
  aes(x = x_het_nl, y = y_het_nl)
) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    linewidth = 0.9,
    colour = "grey30",
    fill = "grey75"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.5,
    colour = "black"
  ) +
  labs(
    x = "Predictor",
    y = "Response"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )


print(plot1_het_nl)

# Plot 2: Residuals vs Fitted values
plot2_het_nl <- ggplot(
  data.frame(
    Fitted    = pred_y_het_nl,
    Residuals = residuals_het_nl
  ),
  aes(x = Fitted, y = Residuals)
) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.55,
    colour = "black"
  ) +
  labs(
    x = "Fitted values",
    y = "Residuals"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )

print(plot2_het_nl)

# TEST
bayesian_homoscedasticity_test(formula = y_het_nl ~ x_het_nl, data = df_het_nl)
```

### Mixed Pattern: Linear and Nonlinear Heteroscedasticity
```{r - Mixed Pattern}
set.seed(123)
n <- 150
x_hybrid <- runif(n, 0, 10)

# Nonlinear component decreasing as x increases,
# linear component becomes more dominant as x increases
sigma_hybrid <- 1 +
  0.20 * (x_hybrid - mean(x_hybrid))^2 +
  0.25 * x_hybrid

y_hybrid <- 1 + 2 * x_hybrid + rnorm(n, sd = sigma_hybrid)

df_hybrid <- data.frame(
  x_hybrid = x_hybrid,
  y_hybrid = y_hybrid,
  sigma_hybrid = sigma_hybrid
)

# Fit mean model
lm_hybrid <- lm(y_hybrid ~ x_hybrid, data = df_hybrid)
pred_y_hybrid <- predict(lm_hybrid)
residuals_hybrid <- resid(lm_hybrid)

# Plot 1: variance structure
plot1_hybrid <- ggplot(
  df_hybrid,
  aes(x = x_hybrid, y = y_hybrid)
) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    linewidth = 0.9,
    colour = "grey30"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.5,
    colour = "black"
  ) +
  labs(
    x = "Predictor",
    y = "Response"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )

print(plot1_hybrid)

# Plot 2: residuals vs fitted values
plot2_hybrid <- ggplot(
  data.frame(
    Fitted = pred_y_hybrid,
    Residuals = residuals_hybrid
  ),
  aes(x = Fitted, y = Residuals)
) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.55,
    colour = "black"
  ) +
  labs(
    x = "Fitted values",
    y = "Residuals"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )

print(plot2_hybrid)

# TEST
bayesian_homoscedasticity_test(formula = y_hybrid ~ x_hybrid, data = df_hybrid)
```

## 3.2 Simulation Study: Performance Across Levels of Heteroscedasticity
### 3.2.1 Sensitivity of the Test to Increasing Levels of Linear Heteroscedasticity 
```{r - Linear Heteroscedasticity}
set.seed(13)
n      <- 220
slopes <- seq(0, 1, length.out = 10)
x_base <- seq(-5, 5, length.out = n)
z_base <- rnorm(n)
all_df <- vector("list", length(slopes))
P1     <- numeric(length(slopes))

for (i in seq_along(slopes)) {
  b      <- slopes[i]
  x      <- x_base
  sigma  <- 4.5 + 0.7 * b * x
  y      <- 1 + 2 * x + sigma * z_base
  df     <- data.frame(x = x, y = y)

  fit <- lm(y ~ x, data = df)
  all_df[[i]] <- data.frame(
    Fitted    = predict(fit),
    Residuals = resid(fit),
    slope     = paste0("b = ", round(b, 2))
  )

  out   <- bayesian_homoscedasticity_test(y ~ x, data = df)
  P1[i] <- out$posterior_prob_H1

  cat(sprintf(" b = %.2f\n", b))
  cat(sprintf("H0: %.3f   H1: %.3f\n\n",
              out$posterior_prob_H0,
              out$posterior_prob_H1))
}

df_all <- do.call(rbind, all_df)
df_res <- data.frame(slope = slopes, Pr_H1 = P1)

p1 <- ggplot(df_all, aes(x = Fitted, y = Residuals)) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  geom_point(
    shape = 16,
    size = 1.5,
    alpha = 0.55,
    colour = "black"
  ) +
  facet_wrap(~ slope, ncol = 5, scales = "free_x") +
  labs(
    x = "Fitted values",
    y = "Residuals"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 13) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.0, "pt"),

    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 11),

    panel.spacing = unit(9, "pt"),
    panel.grid = element_blank()
  )

print(p1)

p2 <- ggplot(df_res, aes(x = slope, y = Pr_H1)) +
  geom_line(linewidth = 0.9, colour = "grey30") +
  geom_point(size = 2.0, colour = "black") +
  labs(
    x =  expression(
      "Strength of linear heteroscedasticity ("* b *")"
     ),
    y = expression(Pr(H[1]))
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0.01, 0.03))
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )

print(p2)
```

### 3.2.2 Sensitivity of the Test to Increasing Levels of Nonlinear Heteroscedasticity
```{r - Nonlinear Heteroscedasticity}
simulate_hetero2 <- function(n = 100, beta0 = 1, beta1 = 2,
                             sigma_base = 1, alpha = 0, k = 2) {
  x <- runif(n, 0, 10)
  mu <- beta0 + beta1 * x
  sigma <- sigma_base * (1 + alpha * (x - mean(x))^k)
  y <- rnorm(n, mean = mu, sd = sigma)
  data.frame(y = y, x = x)
}

set.seed(23)
n_sim  <- 10
alphas <- seq(0, 1.0, length.out = n_sim)
all_df <- vector("list", length(n_sim))
P      <- numeric(length(n_sim))  

for (i in seq_along(alphas)) {
  a  <- alphas[i]
  df <- simulate_hetero2(n = 100, alpha = a)

  fit <- lm(y ~ x, data = df)
  all_df[[i]] <- data.frame(
    Fitted    = predict(fit),
    Residuals = resid(fit),
    alpha     = paste0("alpha = ", round(a, 2))
  )

  out <- bayesian_homoscedasticity_test(y ~ x, data = df)
  P[i] <- out$posterior_prob_H2

  cat(sprintf(" alpha = %.2f\n", a))
  cat(sprintf("H0: %.3f   H2: %.3f\n\n",
              out$posterior_prob_H0,
              out$posterior_prob_H2))
}

df_all <- do.call(rbind, all_df)
df_res <- data.frame(alpha = alphas, Pr_H2 = P)


p1 <- ggplot(df_all, aes(x = Fitted, y = Residuals)) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  geom_point(
    shape = 16,
    size = 1.5,
    alpha = 0.55,
    colour = "black"
  ) +
  facet_wrap(~ alpha, ncol = 5, scales = "free_x") +
  labs(
    x = "Fitted values",
    y = "Residuals"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 13) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.0, "pt"),

    strip.background = element_blank(),

    strip.text = element_text(face = "bold", size = 11),

    panel.spacing = unit(9, "pt"),
    panel.grid = element_blank()
  )

print(p1)

p2 <- ggplot(df_res, aes(x = alpha, y = Pr_H2)) +
  geom_line(
    linewidth = 0.9,
    colour = "grey30"
  ) +
  geom_point(
    size = 2.0,
    colour = "black"
  ) +
  labs(
    x = expression(
      "Strength of quadratic heterosedasticity ("*alpha*")"
    ),
    y = expression(Pr(H[2]))
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0.01, 0.03))
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )


print(p2)
```

## 3.3 Simulation Studies: Sample Size Impact 
### 3.3.1 Linear Heteroscedasticity
```{r - Linear Heteroscedasticity}
set.seed(450)
sample_sizes <- c(40, 100, 200, 500, 1000)
b <- 0.8

Pr_H1 <- numeric(length(sample_sizes))

for (i in seq_along(sample_sizes)) {
  n <- sample_sizes[i]
  x <- runif(n, -5, 5)
  sigma <- 4 + 0.6 * b * x
  y <- 1 + 2 * x + rnorm(n, sd = sigma)
  df <- data.frame(x = x, y = y)

  out <- bayesian_homoscedasticity_test(y ~ x, data = df)

  cat(sprintf(" n = %d\n", n))
  cat(sprintf("H0: %.3f   H1: %.3f\n\n",
              out$posterior_prob_H0,
              out$posterior_prob_H1))

  Pr_H1[i] <- out$posterior_prob_H1
}

df_plot <- data.frame(n = sample_sizes, Pr_H1 = Pr_H1)

p_n <- ggplot(df_plot, aes(x = n, y = Pr_H1)) +
  geom_line(
    linewidth = 0.9,
    colour = "grey30"
  ) +
  geom_point(
    size = 2.0,
    colour = "black"
  ) +
  scale_x_continuous(
    breaks = sample_sizes,
    expand = expansion(mult = c(0.02, 0.04))
  ) +
  scale_y_continuous(
    limits = c(0.5, 1),
    breaks = seq(0.5, 1, 0.1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    x = "Sample size (n)",
    y = expression(Pr(H[1]))
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )
print(p_n)
```

### 3.3.2 Nonlinear Heteroscedasticity
```{r - Nonlinear Heteroscedasticity}
set.seed(12)
sample_sizes <- c(45, 100, 200, 500, 1000)
b <- 0.08

Pr_H2 <- numeric(length(sample_sizes))

for (i in seq_along(sample_sizes)) {
  n <- sample_sizes[i]
  x <- runif(n, 0, 10)
  sigma <- 1 + b * (x - mean(x))^2
  y <- 1 + 2 * x + rnorm(n, sd = sigma)
  df <- data.frame(x = x, y = y)

  out <- bayesian_homoscedasticity_test(y ~ x, data = df)

  cat(sprintf(" n = %d\n", n))
  cat(sprintf("H0: %.3f   H2: %.3f\n\n",
              out$posterior_prob_H0,
              out$posterior_prob_H2))

  Pr_H2[i] <- out$posterior_prob_H2
}

df_plot <- data.frame(n = sample_sizes, Pr_H2 = Pr_H2)

p_n_quad <- ggplot(df_plot, aes(x = n, y = Pr_H2)) +
  geom_line(
    linewidth = 0.9,
    colour = "grey30"
  ) +
  geom_point(
    size = 2.0,
    colour = "black"
  ) +
  scale_x_continuous(
    breaks = sample_sizes,
    expand = expansion(mult = c(0.02, 0.04))
  ) +
  scale_y_continuous(
    limits = c(0.1, 1),
    breaks = seq(0.1, 1, 0.1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    x = "Sample size (n)",
    y = expression(Pr(H[2]))
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line  = element_line(linewidth = 0.8),
    axis.ticks = element_line(linewidth = 0.8),
    axis.ticks.length = unit(3.5, "pt"),
    panel.grid = element_blank()
  )

print(p_n_quad)
```

## 3.4 Prior Sensitivity Analysis: Model Priors
```{r - All Three Hypotheses}
set.seed(120)

n <- 120
x <- runif(n, -4, 4)
x_center <- x - mean(x)
x_shift <- x - min(x)

df_homo <- data.frame(
  y = 2 + 3*x + rnorm(n, 0, 1),
  x = x
)

df_lin <- data.frame(
  y = 2 + 3*x + rnorm(n, 0, 0.6 + 0.20*x_shift),
  x = x
)

df_quad <- data.frame(
  y = 2 + 3*x + rnorm(n, 0, 0.6 + 0.24*x_center^2),
  x = x
)

datasets <- list(
  "Homoscedastic" = df_homo,
  "Linear heteroscedasticity" = df_lin,
  "Quadratic heteroscedasticity" = df_quad
)

prior_vals <- seq(0.1, 0.9, length.out = 9)

results_all <- list()

for (scenario in names(datasets)) {

  df <- datasets[[scenario]]
  results <- data.frame()

  for (pH0 in prior_vals) {

    priors <- c(
      H0 = pH0,
      H1 = (1 - pH0) / 2,
      H2 = (1 - pH0) / 2
    )

    test_res <- bayesian_homoscedasticity_test(
      y ~ x,
      data = df,
      priors = priors,
      account_uncertainty = FALSE
    )

    results <- rbind(results, data.frame(
      prior_H0 = pH0,
      post_H0 = test_res$posterior_prob_H0,
      post_H1 = test_res$posterior_prob_H1,
      post_H2 = test_res$posterior_prob_H2
    ))
  }

  results_long <- results %>%
    pivot_longer(
      cols = c(post_H0, post_H1, post_H2),
      names_to = "posterior",
      values_to = "value"
    ) %>%
    mutate(
      posterior = factor(
        posterior,
        levels = c("post_H0", "post_H1", "post_H2"),
        labels = c("H0", "H1", "H2")
      )
    )

  p <- ggplot(
    results_long,
    aes(
      x = prior_H0,
      y = value,
      linetype = posterior,
      shape = posterior
    )
  ) +
    geom_line(linewidth = 0.9, colour = "grey30") +
    geom_point(size = 2.0, colour = "black") +
    scale_linetype_manual(values = c(H0 = "solid", H1 = "22", H2 = "42")) +
    scale_shape_manual(values = c(H0 = 16, H1 = 17, H2 = 15)) +
    scale_x_continuous(
      breaks = prior_vals,
      limits = c(0.1, 0.9),
      expand = expansion(mult = c(0.02, 0.04))
    ) +
    scale_y_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, 0.2),
      expand = expansion(mult = c(0.05, 0.03))
    ) +
    labs(
      x = expression("Prior probability " * Pr(H[0])),
      y = "Posterior probability",
      linetype = NULL,
      shape = NULL,
      subtitle = scenario
    ) +
    theme_classic(base_size = 14) +
    theme(
      axis.line = element_line(linewidth = 0.8),
      axis.ticks = element_line(linewidth = 0.8),
      axis.ticks.length = unit(3.5, "pt"),
      panel.grid = element_blank(),
      legend.position = "top",
      legend.key.width = unit(18, "pt"),
      plot.subtitle = element_text(face = "bold", margin = margin(b = 6))
    )

  print(p)

  results_all[[scenario]] <- results
}
```

## 3.5  Prior Sensitivity Analysis: Parameter Priors
```{r - All Three Hypotheses}
set.seed(30)
n <- 100
x <- runif(n, 0, 10)
x_center <- x - mean(x)

df_homo <- data.frame(
  y = 2 + 3*x + rnorm(n, 0, 1),
  x = x
)

df_lin <- data.frame(
  y = 2 + 3*x + rnorm(n, 0, 0.5 + 0.2*x),
  x = x
)

df_quad <- data.frame(
  y = 2 + 3*x + rnorm(n, 0, 0.6 + 0.22*x_center^2),
  x = x
)

datasets <- list(
  "H0: Homoscedastic" = df_homo,
  "H1: Linear heteroscedastic" = df_lin,
  "H2: Quadratic heteroscedastic" = df_quad
)

jzs_scales <- c("medium", "wide", "ultrawide")
priors <- c(H0 = 1/3, H1 = 1/3, H2 = 1/3)

res_all <- list()

for (nm in names(datasets)) {

  df <- datasets[[nm]]

  tmp_jzs <- lapply(jzs_scales, function(rsc) {

    out <- bayesian_homoscedasticity_test(
      y ~ x,
      data = df,
      priors = priors,
      evidence = "JZS",
      rscale = rsc,
      account_uncertainty = FALSE
    )

    data.frame(
      scenario = nm,
      method = paste0("JZS (", rsc, ")"),
      Pr_target = if (startsWith(nm, "H0")) {
        out$posterior_prob_H0
      } else if (startsWith(nm, "H1")) {
        out$posterior_prob_H1
      } else {
        out$posterior_prob_H2
      }
    )
  })

  out_bic <- bayesian_homoscedasticity_test(
    y ~ x,
    data = df,
    priors = priors,
    evidence = "BIC",
    account_uncertainty = FALSE
  )

  tmp_bic <- data.frame(
    scenario = nm,
    method = "BIC",
    Pr_target = if (startsWith(nm, "H0")) {
      out_bic$posterior_prob_H0
    } else if (startsWith(nm, "H1")) {
      out_bic$posterior_prob_H1
    } else {
      out_bic$posterior_prob_H2
    }
  )

  res_all[[nm]] <- bind_rows(tmp_jzs, tmp_bic)
}

df_prior_sens <- bind_rows(res_all)

df_prior_sens$method <- factor(
  df_prior_sens$method,
  levels = c(
    "JZS (medium)",
    "JZS (wide)",
    "JZS (ultrawide)",
    "BIC"
  )
)

plot_one_scenario <- function(df_plot, scenario_name) {

  ggplot(
    df_plot %>% filter(scenario == scenario_name),
    aes(x = method, y = Pr_target, group = 1)
  ) +
    geom_line(linewidth = 0.9, colour = "grey30") +
    geom_point(size = 2.0, colour = "black") +
    scale_y_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, 0.2),
      expand = expansion(mult = c(0.05, 0.03))
    ) +
    labs(
      title = scenario_name,
      x = "Evidence specification",
      y = "Posterior probability of target model"
    ) +
    theme_classic(base_size = 14) +
    theme(
      axis.line = element_line(linewidth = 0.8),
      axis.ticks = element_line(linewidth = 0.8),
      axis.ticks.length = unit(3.5, "pt"),
      axis.text.x = element_text(angle = 15, hjust = 1),
      plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
      panel.grid = element_blank()
    )
}

p_homo <- plot_one_scenario(df_prior_sens, "H0: Homoscedastic")
p_lin  <- plot_one_scenario(df_prior_sens, "H1: Linear heteroscedastic")
p_quad <- plot_one_scenario(df_prior_sens, "H2: Nonlinear heteroscedastic")

print(p_homo)
print(p_lin)
print(p_quad)
```


## 3.6 Application to Real-Life Data
### 3.6.1 Homoscedastic Data
```{r - Homoscedastic Data}
data("survey", package = "MASS")
data_h0 <- na.omit(survey[, c("NW.Hnd", "Wr.Hnd")])

# Fit linear model
model_h0 <- lm(NW.Hnd ~ Wr.Hnd, data = data_h0)

# Add fitted values and residuals to dataset
data_h0$Fitted <- fitted(model_h0)
data_h0$Residuals <- resid(model_h0)

# 1. Variance Structure Plot
p1_h0 <- ggplot(data_h0, aes(x = Wr.Hnd, y = NW.Hnd)) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    linewidth = 0.9,
    colour = "grey30"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.5,
    colour = "black"
  ) +
  labs(
    x = "Writing hand span",
    y = "Non-writing hand span"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(panel.grid = element_blank())

p1_h0


# 2. Residuals vs Fitted Values
p2_h0 <- ggplot(data_h0, aes(x = Fitted, y = Residuals)) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.55,
    colour = "black"
  ) +
  labs(
    x = "Fitted values",
    y = "Residuals"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(panel.grid = element_blank())

p2_h0

# BP test (linear heteroscedasticity)
bptest(model_h0)

# White test (nonlinear heteroscedasticity)
whitestrap::white_test(model_h0)

# Bayesian test
bayesian_homoscedasticity_test(NW.Hnd ~ Wr.Hnd, data = data_h0)
```

### 3.6.2 Linear Heteroscedastic Data
```{r - Linear Heteroscedastic Data}
data("crabs", package = "MASS")
data <- crabs

fit_crabs <- lm(RW ~ CW, data = data)
data$Fitted    <- predict(fit_crabs)
data$Residuals <- resid(fit_crabs)

p1_crabs <- ggplot(data, aes(x = CW, y = RW)) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    linewidth = 0.9,
    colour = "grey30"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.5,
    colour = "black"
  ) +
  labs(
    x = "Carapace width",
    y = "Rear width"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(panel.grid = element_blank())

p1_crabs

p2_crabs <- ggplot(data, aes(x = Fitted, y = Residuals)) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.55,
    colour = "black"
  ) +
  labs(
    x = "Fitted values",
    y = "Residuals"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(panel.grid = element_blank())

p2_crabs

fit <- lm(RW ~ CW, data = data)

# BP test (linear heteroscedasticity)
bptest(fit)

# White test (nonlinear heteroscedasticity)
whitestrap::white_test(fit)

# Bayesian test
bayesian_homoscedasticity_test(RW ~ CW, data = data)
```

### 3.6.3 Nonlinear Heteroscedastic Data
```{r - Nonlinear Heteroscedastic Data}
data("engel", package = "quantreg")
data_t <- engel
head(data_t)

model_man <- lm(foodexp ~ income, data = data_t)

data_t$Fitted <- fitted(model_man)
data_t$Residuals <- resid(model_man)

# 1. Variance Structure Plot
p1_engel <- ggplot(data_t, aes(x = income, y = foodexp)) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    linewidth = 0.9,
    colour = "grey30"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.5,
    colour = "black"
  ) +
  labs(
    x = "Income",
    y = "Food expenditure"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(panel.grid = element_blank())

p1_engel

# 2. Residuals vs Fitted Values
p2_engel <- ggplot(data_t, aes(x = Fitted, y = Residuals)) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.7,
    colour = "grey35"
  ) +
  geom_point(
    shape = 16,
    size = 1.6,
    alpha = 0.55,
    colour = "black"
  ) +
  labs(
    x = "Fitted values",
    y = "Residuals"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.04))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.08))) +
  theme_classic(base_size = 14) +
  theme(panel.grid = element_blank())

p2_engel

# BP test (linear heteroscedasticity)
bptest(model_man)

# White test (nonlinear heteroscedasticity)
whitestrap::white_test(model_man)

# Bayesian test
bayesian_homoscedasticity_test(foodexp ~ income, data = data_t)
```

## 3.7 Extended Application: Accounting for Residual Uncertainty
```{r - Homoscedasticity}
bayesian_homoscedasticity_test(formula = y_homosc ~ x_homosc, data = df_homosc, account_uncertainty = TRUE)
```

```{r - Linear Heteroscedasticity}
bayesian_homoscedasticity_test(formula = y_het_lin ~ x_het_lin, data = df_het_lin, account_uncertainty = TRUE)
```

```{r - Nonlinear Heteroscedasticity}
bayesian_homoscedasticity_test(formula = y_het_nl ~ x_het_nl, data = df_het_nl, account_uncertainty = TRUE)
```

